Peaks Over Threshold for Bursty Time Series

### 2018-04-17 article_springer/article_springer.Rmd
likelihood ratio test

However, as described nicely by @Resnick97, the less 
similar a heavy-tailed distribution is to a Pareto distribution, the less
useful a Hill estimator is. 
The top-right panel of \ref{fig:QQ-plots}, for instance, shows the 
QQ-estimator plot for $2000$ draws from the Mittag-Leffler distribution with 
tail parameter $0.8$. The stretched exponential shape means 
that the Mittag-Leffler distribution has more probability near $0$ than 
the Pareto distribution, severely biasing the estimator downwards 
(`r round(ML_exponent, 2)`). Hence a cutoff needs to be introduced, 
above which the tail exponent is fitted: The dynamic QQ-estimator 
plots the static QQ-estimates (y-axis) for different cutoffs (x-axis)
(see bottom-left panel). There is no apparent interval of stability, 
which could be taken to indicate a reasonable estimate for the true 
tail exponent $0.8$. 
Even worse, from a QQ-estimator point of view, there is no striking 
difference between the infinite mean ${\rm ML}(0.8,1)$ distribution and 
the light-tailed ${\rm Exp}(1)$ distribution (see bottom-right panel). 
In particular, if only few 
(e.g. $<100$) data are available, then the static QQ-estimator plots look 
virtually interchangeable.

Hence if there is some prior expectation that the data are drawn from the 
Mittag-Leffler distribution
(as is the the case for threshold exceedance times),
then we recommend avoiding QQ-estimators and Hill plots altogether, 
and instead examining QQ-plots directly 
(see Figure \ref{fig:QQ-plots}). 
The scale parameter $\sigma$ is irrelevant for a QQ-Plot. 
The tail parameter $\beta$ can be estimated quickly via the log-moment 
method by @Cahoy2013, or via maximum likelihood. 
Both estimators are implemented in the R package `MittagLeffleR` 
[@MittagLeffleR].

```{r QQ-plots, fig.cap="QQ Plots for Mittag-Leffler data (left; $\\beta = 0.8$) and threshold crossing times / magnitudes for the solar flare data (middle / right; threshold is set at the 500th order statistic, $\\beta = 0.83$). Threshold crossing times appear to fit the Mittag-Leffler distribution well, whereas the Magnitudes appear to fit a Pareto distribution better (\\ref{fig:QQ-est}). "}
par(mfrow = c(1, 3))
rml(n = 500, tail = 0.8) %>% mlqqplot(tail = 0.8, log = 'xy', main = "Mittag-Leffler data")
flares_arrivals <- flares %>% ctrm() %>% thin(k = 500) %>% interarrival()
flares_arrivals_MLparam <- (flares_arrivals %>% mlmle())$par
mlqqplot(flares_arrivals, 
         tail = flares_arrivals_MLparam["tail"], 
         scale = flares_arrivals_MLparam["scale"],
         log = 'xy', 
         main = "Threshold crossing times")
flares %>% ctrm() %>% thin(k = 500) %>% coredata() %>% 
  mlqqplot(tail = 0.8, log = 'xy', main = "Magnitudes")
```

>>>>

However, as described nicely by @Resnick97, the less 
similar a heavy-tailed distribution is to a Pareto distribution, the less
useful a Hill estimator is. 
The top-right panel of \ref{fig:QQ-plots}, for instance, shows the 
QQ-estimator plot for $2000$ draws from the Mittag-Leffler distribution with 
tail parameter $0.8$. The stretched exponential shape means 
that the Mittag-Leffler distribution has more probability near $0$ than 
the Pareto distribution, severely biasing the estimator downwards 
(`r round(ML_exponent, 2)`). Hence a cutoff needs to be introduced, 
above which the tail exponent is fitted: The dynamic QQ-estimator 
plots the static QQ-estimates (y-axis) for different cutoffs (x-axis)
(see bottom-left panel). There is no apparent interval of stability, 
which could be taken to indicate a reasonable estimate for the true 
tail exponent $0.8$. 
Even worse, from a QQ-estimator point of view, there is no striking 
difference between the infinite mean ${\rm ML}(0.8,1)$ distribution and 
the light-tailed ${\rm Exp}(1)$ distribution (compare the right two panels). 
In particular, if only few (e.g. $<100$) data are available, 
then the static QQ-estimator plots look interchangeable for the two 
distributions.

```{r likelihood-ratio-test}
flares_arrivals <- flares %>% ctrm() %>% thin(k = 500) %>% interarrival()
mlmle_out <- mlmle(data = flares_arrivals)
```

Hence if there is some prior expectation that the data are drawn from the 
Mittag-Leffler distribution
(as is the the case for threshold exceedance times),
then we recommend avoiding QQ-estimators and Hill plots altogether, 
and instead examining QQ-plots directly 
(see Figure \ref{fig:QQ-plots}). 
The scale parameter $\sigma$ is irrelevant for a QQ-Plot. 
The tail parameter $\beta$ can be estimated quickly via the log-moment 
method by @Cahoy2013, or via maximum likelihood. 
Both estimators are implemented in the R package `MittagLeffleR` 
[@MittagLeffleR]. 
Since the exponential distribution is nested in the Mittag-Leffler 
family of distributions, a standard likelihood ratio test can be 
performed, with the exponential distribution as a null model against 
a Mittag-Leffler distribution as the alternative. 
As an example, the threshold crossing times for the solar flare dataset
yield a difference in deviance of 
$\approx$ `r round(2 * (mlmle_out$loglik - mlmle_out$null_loglik))`, 
which evaluates to a strongly significant $\chi^2_1$ test statistic 
(p-value = $10^{`r round(log10(mlmle_out$p_value))`}$) for the null
hypothesis $\beta = 1$.

```{r QQ-plots, fig.cap="QQ Plots for Mittag-Leffler data (left; $\\beta = 0.8$) and threshold crossing times / magnitudes for the solar flare data (middle / right; threshold is set at the 500th order statistic, $\\beta = 0.83$). Threshold crossing times appear to fit the Mittag-Leffler distribution well, whereas the Magnitudes appear to fit a Pareto distribution better (\\ref{fig:QQ-est}). "}
par(mfrow = c(1, 3))
rml(n = 500, tail = 0.8) %>% mlqqplot(tail = 0.8, log = 'xy', main = "Mittag-Leffler data")
flares_arrivals_MLparam <- (flares_arrivals %>% mlmle())$par
mlqqplot(flares_arrivals, 
         tail = flares_arrivals_MLparam["tail"], 
         scale = flares_arrivals_MLparam["scale"],
         log = 'xy', 
         main = "Threshold crossing times")
flares %>% ctrm() %>% thin(k = 500) %>% coredata() %>% 
  mlqqplot(tail = 0.8, log = 'xy', main = "Magnitudes")
```
